---
output: html_document
editor_options: 
 chunk_output_type: console
---

# Get and save Congressional District weights

Get a wide tibble (data frame) with one row per tmd record and a column for the 2021 weight of each of the 435 Congressional districts plus DC (column names are ak00, ..., wy00). Save it in the `intermediate` folder of this project. The file is approximately 380 mb.

This needs to be done each time new CD weights are created. The weights files need to be copied into the `temp_data` folder of this R project from the folder in which they were created.

This program is time consuming (~2+ minutes) when weights for all 436 districts are read, so set `eval: false` in the `get-save-weights` chunk after an initial run that gets the data.

## Setup

```{r}
#| label: setup
#| output: false

source(here::here("R", "libraries.R"))
source(here::here("R", "constants.R"))

```

## Get and save weights

Get and save weights each time new weights are produced. Set `eval: false` in the `get-save-weights` chunk after weights are updated.

```{r}
#| label: get-save-weights
#| eval: false
#| output: false

weightfiles <- dir_ls(here::here("temp_data")) |> str_subset("tmd_weights.csv.gz")

f <- function(fpath){
  cd <- str_sub(fs::path_file(fpath), 1, 4)
  print(cd)
  vroom(fpath, col_select = "WT2021") |> 
    rename(!!cd := WT2021)
}

a <- proc.time()
weights <- test <- purrr::map(weightfiles, f) |> 
  list_cbind() |> 
  mutate(row=row_number()) |> 
  relocate(row)
b <- proc.time()
b - a # 79 seconds

system.time(saveRDS(weights, here::here("intermediate", "weights.rds"))) # 23 seconds

# system.time(cdweights <- readRDS(here::here("intermediate", "weights.rds"))) # 4.5 secs, 390 mb

```


