---
output: html_document
editor_options: 
 chunk_output_type: console
---

# Get and save selected ACS data

Get median household income for each 118th-session Congressional District in 2022. The original source is the 2022 Census 1-year American Community Survey (ACS). May be useful in understanding some differences across CDs.

This only needs to be done once so set `eval: false` in the `get-save-acs` chunk after getting the data.

## Setup

```{r}
#| label: setup
#| output: false

source(here::here("R", "system_specific_constants.R"))

source(here::here("R", "libraries.R"))
source(here::here("R", "constants.R"))
source(here::here("R", "functions.R"))

```


## ONETIME download of data

```{r}
#| label: get-save-acs
#| eval: false
#| output: false

mhhi2022 <- get_acs(
  geography = "congressional district",
  variables = "S1903_C03_001",  # Median household income variable S1903_C03_001E
  year = 2022,
  survey = "acs1"  # Consider using 5-year estimates for better coverage, especially in smaller areas
) # 118th Congress

# mhhi2021 <- get_acs(
#   geography = "congressional district",
#   variables = "S1903_C03_001",  # Median household income variable S1903_C03_001E
#   year = 2021,
#   survey = "acs1"  # Consider using 5-year estimates for better coverage, especially in smaller areas
# ) # 116th Congress

stcodes <- tigris::states() |> 
  as_tibble() |> 
  select(stfips=GEOID, stabbr=STUSPS, stname=NAME) |> 
  arrange(stfips)

# add our codes
mhhi2022a <- mhhi2022 |> 
  mutate(stfips=str_sub(GEOID, 1, 2),
         cd=str_sub(GEOID, 3, 4)) |> 
  left_join(stcodes, by = join_by(stfips)) |> 
  mutate(statecd=paste0(str_to_lower(stabbr), cd),
         session=118)

count(mhhi2022a, stfips, stabbr, stname) |> 
  janitor::adorn_totals() # 435 CDs + DC + PR = 437 records

mhhi2022b <- mhhi2022a |> 
  select(session, stname, statecd, mhhi=estimate)

write_csv(mhhi2022b, here::here("raw_data", "mhhi2022.csv"))

```
