---
output: html_document
editor_options: 
 chunk_output_type: console
execute:
  freeze: false
 
---

# Congressional District analysis -- individual district reports

```{r}
#| label: setup
#| output: false

source(here::here("R", "libraries.R"))
source(here::here("R", "functions.R"))
source(here::here("R", "functions_constants.R"))

```


```{r}
#| label: constants
#| output: false

AREA_TYPE <- "cd" # if want to redefine constants
# AREA_TYPE <- "test" # if want to redefine constants

CONSTANTS <- get_constants(AREA_TYPE)
# CONSTANTS
# CONSTANTS$WEIGHTS_DIR

```

## Explanation of individual-district tables

The next sections have results for selected states, in filterable and searchable form.

Units:

-   Dollar amounts are in $ millions (varname==amount for target, wtdvalue, and diff)
-   Counts (including XTOT) are actual numbers

scope:

-   0 = total population
-   1 = filers
-   2 = nonfilers (none currently in the table)

fstatus:

-   0 = sum of all statuses
-   1 = married joint
-   2 = single
-   3 = married filing separately (not targeted)
-   4 = head of household

targeted:

-  true = item was targeted
-  false = variable was in target file but item was not targeted (e.g., amount was targeted but not count, or item was commented out)

Dropdown boxes and search fields allow narrowing down the records that are displayed.


```{r}
#| results: asis
#| echo: false

compfile <- readr::read_csv(fs::path(CONSTANTS$OUTPUT_DIR, "compfile.csv"),
                                             show_col_types = FALSE)

# areas <- c(CONSTANTS$PHASE6_STATES, "SC") |> stringr::str_to_lower()
areas <- c("AK00", "DE00", "ID01", "ID02", "ME02", "MT00", "ND00", "PA08", "SD00", "WY00") |> stringr::str_to_lower()
# areas <- c("ak")

for(area in areas) {
    template <- readLines("_area-template_dt.qmd")
    template <- str_replace_all(template, coll("{{area}}"), area)
    cat(knitr::knit_child(text = template, quiet = TRUE))
    cat("\n\n")
}

```
