---
params:
  area_type: "placeholder"
  eval_data: FALSE
engine: knitr  
output: html_document
editor_options: 
 chunk_output_type: console
---

# Data preparation

```{r}
#| label: setup
#| output: false

source(here::here("R", "libraries.R"))
source(here::here("R", "functions.R"))
source(here::here("R", "functions_constants.R"))

```


```{r}
#| label: constants
#| output: false

AREA_TYPE = params$area_type

CONSTANTS <- get_constants(AREA_TYPE)
CONSTANTS

```

Prepare data for `r CONSTANTS$LONG_NAME`.

## ONETIME: Get and save weights

Get and save a wide tibble of weights each time new weights are produced. Once the file is created and saved, retrieval is much faster.

```{r}
#| label: ONETIME-save-weights
#| eval: !expr params$eval_data
#| output: false

save_weights()

```

## ONETIME: Supplement the weights file with selected TMD variables and summarize

Supplement the weights with selected TMD variables and summarize. Update when area weights change, or the national TMD data change, or if we choose additional or alternative TMD variables to summarize.


```{r}
#| label: ONETIME-save-enhanced-weighted-tmd-sums
#| eval: !expr params$eval_data
#| output: false

TAXCALC_VARS <- c("XTOT", "c00100", "e00200", "e00300", "e01500", "e02400", "e26270", "e18400", "e18500", "iitax")
save_enhanced_weighted_sums(TAXCALC_VARS)

```

## Prepare target comparison file

```{r}
#| label: save-final-file
#| eval: !expr params$eval_data
#| output: false

compfile <- get_combined_file()
write_csv(compfile, fs::path(CONSTANTS$OUTPUT_DIR, "compfile.csv"))
# skim_without_charts(compfile)

```


