---
params:
  area_type: "placeholder"
  eval_data: FALSE
engine: knitr  
output: html_document
editor_options: 
 chunk_output_type: console
---

# Individual area reports

```{r}
#| label: setup
#| output: false

source(here::here("R", "libraries.R"))
source(here::here("R", "functions.R"))
source(here::here("R", "functions_constants.R"))

```


```{r}
#| label: constants
#| output: false

AREA_TYPE = params$area_type

CONSTANTS <- get_constants(AREA_TYPE)
# CONSTANTS

```

Reports for selected individual `r CONSTANTS$LONG_NAME`.

## Explanation of tables for individual `r CONSTANTS$LONG_NAME`

The next sections have results for selected areas, in filterable and searchable form.

Units:

-   Dollar amounts are in $ millions (varname==amount for target, wtdvalue, and diff)
-   Counts (including XTOT) are actual numbers

scope:

-   0 = total population
-   1 = filers
-   2 = nonfilers (none currently in the table)

fstatus:

-   0 = sum of all statuses
-   1 = married joint
-   2 = single
-   3 = married filing separately (not targeted)
-   4 = head of household

targeted:

-  true = item was targeted
-  false = variable was in target file but item was not targeted (e.g., amount was targeted but not count, or item was commented out)

Dropdown boxes and search fields allow narrowing down the records that are displayed.


**Note**: Because we do not have targets for iitax we show only total iitax and number of returns with nonzero iitax. Additional summaries are available in the comparison file created in this R project.


```{r}
#| label: create-area-reports
#| results: asis
#| echo: false

compfile <- readr::read_csv(fs::path(CONSTANTS$OUTPUT_DIR, "compfile.csv"),
                                             show_col_types = FALSE)

for(area in CONSTANTS$AREAS) {
    template <- readLines("_area-template_dt.qmd")
    template <- stringr::str_replace_all(template, coll("{{area}}"), area)
    cat(knitr::knit_child(text = template, quiet = TRUE))
    cat("\n\n")
}

```
