---
output: html_document
editor_options: 
 chunk_output_type: console
---

# Get and save state weights

Get a wide tibble (data frame) with one row per tmd record and a column for the 2021 weight of each area (column names are ak00, ..., wy00). Save it as `cdweights.rds` in the `intermediate` folder of this project. For Congressional Districts the file is approximately 380 mb.

This needs to be done each time new area weights are created. The weights files need to be copied into the `temp_data` folder of this R project from the folder in which they were created.

This program is time consuming (~2+ minutes) when weights for all 436 districts are read, so set `eval: false` in the `get-save-weights` chunk after an initial run that gets the data.

## Setup

```{r}
#| label: setup
#| output: false

source(here::here("R", "libraries.R"))

source(here::here("R", "system_specific_constants.R"))
source(here::here("R", "constants.R"))

source(here::here("R", "functions.R"))

```

## Get and save weights

Get and save weights each time new weights are produced. Set `eval: false` in the `get-save-weights` chunk after weights are updated.

Create a data frame with one row per tax record and one column per area with the value of `WT2021` for that area.

Steps:

-   Read the 2021 weights column from each area weights `_tmd_weights.csv.gz` file in the `temp_data` folder
-   Concatenate
-   Add a `row` (row number) column to make sorting and merging safer when the file is used
-   Save as `cdweights.rds` in the `intermediate`folder

The resulting .rds file is about 390 mb.

This is time consuming. On Boyd's computer reading and concatenating takes about 80-90 seconds, and saving takes about 25 seconds.

Once the file is created and saved, retrieving it only takes 3-4 seconds.

```{r}
#| label: get-save-weights
#| eval: false
#| output: false

# initial WEIGHTSDIR /home/donboyd5/Documents/python_projects/scratch/phase5_salt
# "\\wsl.localhost\Ubuntu-24.04\home\donboyd5\Documents\python_projects\backups\tmd_items_2024-12-20\states"
# pp <- "/home/donboyd5/Documents/python_projects"
# WEIGHTSDIR <- fs::path(pp, "backups/tmd_items_2024-12-20/states")

statedir <- "/mnt/g/.shortcut-targets-by-id/1pEdofaxeQgEeDLM8NOpo0vOGL1jT8Qa1/AFPI_2024/Phase 6/states/"
areatype <- "state" # state or cd

df <- get_weights(statedir, "state")
glimpse(df)

df2 <- readRDS(here::here("intermediate", paste0(areatype, "_weights.rds")))

```
