---
params:
  area_type: "placeholder"
  eval_data: FALSE
engine: knitr  
output: html_document
editor_options: 
 chunk_output_type: console
---

# Variables included in the analysis

```{r}
#| label: setup
#| output: false

source(here::here("R", "libraries.R"))
source(here::here("R", "functions.R"))
source(here::here("R", "functions_constants.R"))

```


```{r}
#| label: constants
#| output: false

AREA_TYPE = params$area_type

CONSTANTS <- get_constants(AREA_TYPE)
CONSTANTS

```


```{r}
#| label: get-data
#| output: false

compfile <- readr::read_csv(fs::path(CONSTANTS$OUTPUT_DIR, "compfile.csv"),
                                             show_col_types = FALSE)

```

The analysis in this report includes variables that were mapped to SOI data, whether targeted or not (see below). It includes the following variables for `r CONSTANTS$LONG_NAME`.

```{r}
#| label: vars-included

vars_included <- compfile |> 
  select(varname, basesoivname, description) |> 
  distinct() 

vars_included |> 
  gt()

```


## What it means for a variable to be mapped to SOI data

When a TMD variable is mapped to SOI data, it means that we were able to develop *potential* area targets for the variable using SOI data. Whether to use those targets in optimization is a separate decision.

A TMD variable is mapped if we have, after careful checking, determined that we can link a specific TMD variable to a specific SOI variable or to a related proxy.

For example:

-   TMD variable c00100 (AGI) links directly to the SOI variable a00100.
-   TMD variable e01500 (Total pension income) does not have a directly corresponding variable in the SOI summary data. However, the SOI data have a01700 (Taxable pension income), which we used as a proxy for linking. Each area's potential target is estimated by sharing total TMD value for e01500 (i.e., its weighted sum for the nation) to individual areas according to its share of the national sum of SOI a01700. (This is done separately for each AGI range.)

Variables that we have linked to the SOI in one of these ways are included in the analysis below. 

It is possible to extend the analysis to additional variables -- and to target additional variables -- but each requires careful checking to determine (1) whether an SOI variable is highly similar in concept to a TMD variable by checking IRS forms and other resources, and (2) how the magnitude of the SOI variable for the nation compares to the related TMD variable for the nation.

## What it means for an item to be targeted

Whether an item is targeted depends on how we defined targeting for a specific variable as well as our general rules of targeting. When we targeted specific variables, we generally targeted each AGI range separately. Because we targeted all AGI ranges, we did not need to target a total for the variable. This makes the optimization more efficient than if we targeted the total.

In some cases, we excluded one or more income ranges from targeting, usually because the target created challenges and we expected it to be unimportant to overall tax liability. Also, we applied a blanket rule and did not target area-variable-agi combinations where the reported SOI value was zero.

Thus, it is possible to have (1) mapped variables that are untargeted, (2) targeted variables with one or more AGI ranges excluded from targeting for all areas, and (3) targeted variables that have an AGI range untargeted for an individual area if the potential target for that area in that AGI range was zero.


