---
output: html_document
editor_options: 
 chunk_output_type: console
---

# Download SOI Congressional District data and documentation files

SOI Congressional District data and documentation files are included with the project in the "../cds/raw_data" folder and by default, rendering the project will NOT re-download the data.

To re-download data and save to the "../cds/raw_data" folder, set option `eval: true` in the `downloads` chunk below

## File locations on the web

-   [SOI Congressional Districts landing page](https://www.irs.gov/statistics/soi-tax-stats-data-by-congressional-district)

    -   [2021 folder](https://www.irs.gov/statistics/soi-tax-stats-data-by-congressional-district-2021)
        -   [zip archive with all data files](https://www.irs.gov/pub/irs-soi/congressional2021.zip)
        -   [Word documentation](https://www.irs.gov/pub/irs-soi/21incddocguide.docx)

## Setup

```{r}
#| label: setup

source(here::here("R", "libraries.R"))
source(here::here("R", "constants.R"))
source(here::here("R", "functions.R"))

```


## (Optionally) Download SOI CD data and documentation

As noted, by default, the files will NOT be downloaded when the project is rendered.

```{r}
#| label: downloads
#| eval: false

# data
fname <- fs::path_file(CDZIPURL)
download.file(url = CDZIPURL, destfile = fs::path(CDRAW, fname), mode = "wb")

# documentation
fname <- fs::path_file(CDDOCURL)
download.file(url = CDDOCURL, destfile = fs::path(CDRAW, fname), mode = "wb")

```


