---
output: html_document
editor_options: 
 chunk_output_type: console
---

# Download SOI state data and documentation files

SOI state data and documentation files are included with the project in the "../data/raw_data" folder and by default, rendering the project will NOT re-download the data.

To re-download data and save to the "../data/raw_data" folder, set option `eval: true` in the `downloads` chunk below

## File locations on the web

-   [SOI State Historical Table 2 landing page with 2020 and 2021 data](https://www.irs.gov/statistics/soi-tax-stats-historic-table-2)

    -   [2021 csv data](https://www.irs.gov/pub/irs-soi/21in55cmcsv.csv)
    -   [2021 documentation](https://www.irs.gov/pub/irs-soi/21incmdocguide.doc)
    -   [2021 Excel file for all states](https://www.irs.gov/pub/irs-soi/21in55cm.xlsx)
    -   also see [2015-2019 folder](https://www.irs.gov/statistics/soi-tax-stats-historic-table-2-2015-2019)

## Setup

```{r}
#| label: setup

source(here::here("R", "libraries.R"))
source(here::here("R", "constants.R"))
source(here::here("R", "functions.R"))

```


## (Optionally) Download SOI state data and documentation

As noted, by default, the files will NOT be downloaded when the project is rendered. 

**All necessary files will be in the `raw_data` folder on GitHub and should be available when the repo is cloned.**

The necessary files are shown below. For those that have [yy], [yy] are the trailing 2 digits of the year. The project includes files for 2015-2021:

-   [yy]in54cm.xlsx or similar, from SOI: each file has human-readable SOI values for all states for the year.
-   [yy]in54cmcsv.csv or similar, from SOI: file with SOI values for all states for the year.
-   [yy]incmdocguide.doc, from SOI: descriptive documentation, including a table of variables and their descriptions
-   soi_states_variable_documentation.xlsx, hand-created by project team from the .doc files: has a sheet for each year with its documentation table. **NOT** intended to be edited by users.

Change options for the chunk below to `eval: true` to download files.

```{r}
#| label: downloads
#| eval: false

# https://www.irs.gov/pub/irs-soi/21in55cmcsv.csv
# https://www.irs.gov/pub/irs-soi/20in55cmcsv.csv
# https://www.irs.gov/pub/irs-soi/19in55cmcsv.csv
# https://www.irs.gov/pub/irs-soi/18in55cmagi.csv
# https://www.irs.gov/pub/irs-soi/17in54cmcsv.csv
# https://www.irs.gov/pub/irs-soi/16in54cmcsv.csv
# https://www.irs.gov/pub/irs-soi/15in54cmcsv.csv

# documentation
# https://www.irs.gov/pub/irs-soi/21incmdocguide.doc
# https://www.irs.gov/pub/irs-soi/20incmdocguide.doc
# https://www.irs.gov/pub/irs-soi/19incmdocguide.doc
# https://www.irs.gov/pub/irs-soi/18incmdocguide.doc
# https://www.irs.gov/pub/irs-soi/17incmdocguide.doc
# https://www.irs.gov/pub/irs-soi/16incmdocguide.doc
# https://www.irs.gov/pub/irs-soi/15incmdocguide.doc

# all-states spreadsheets
# https://www.irs.gov/pub/irs-soi/21in55cm.xlsx
# https://www.irs.gov/pub/irs-soi/20in55cm.xlsx
# https://www.irs.gov/pub/irs-soi/19in55cm.xlsx
# https://www.irs.gov/pub/irs-soi/18in55cm.xlsx
# https://www.irs.gov/pub/irs-soi/17in54cm.xlsx
# https://www.irs.gov/pub/irs-soi/16in54cm.xlsx
# https://www.irs.gov/pub/irs-soi/15in54cm.xlsx


# download 2015-2021 data
get_data <- function(year){
  soiurl <- r"(https://www.irs.gov/pub/irs-soi/)"
  doc_suffix <- "incmdocguide.doc"
  
  csv_suffix <- case_when(
    year %in% 2015:2017 ~ "in54cmcsv.csv",
    year == 2018 ~ "in55cmagi.csv",
    year %in% 2019:2021 ~ "in55cmcsv.csv")
  
  xlsx_suffix <- case_when(
    year %in% 2015:2017 ~ "in54cm.xlsx",
    year %in% 2018:2021 ~ "in55cm.xlsx")  
  
  dl <- function(y2, suffix){
    fname <- paste0(y2, suffix)
    url <- fs::path(soiurl, fname)
    savedir <- fs::path(here::here(), "data", "data_raw")
    savepath <- fs::path(savedir, fname)
    download.file(url, savepath, mode="wb")
  }
  
  print(paste0("Getting data for: ", year))
  y2 <- str_sub(year, 3, 4)
  dl(y2, csv_suffix)
  dl(y2, doc_suffix)
  dl(y2, xlsx_suffix)
}

purrr::walk(2015:2021, get_data)

```


