---
output: html_document
editor_options: 
 chunk_output_type: console
---

# Create base file for 117th Congressional District target files

This section creates one long file that is a superset of what we need for individual 117th Congressional District target files. This long file has everything needed to extract and save a target file for any CD . It also has additional convenience variables that will not be included in individual CD target files such as variable descriptions, human-friendly AGI-range labels, state fips codes, and a sort code for ordering records within a CD. These variables are excluded when target files are written.

## Setup

```{r}
#| label: setup

source(here::here("R", "libraries.R"))
source(here::here("R", "constants.R"))
source(here::here("R", "functions.R"))

# 334283385.27000004 national pop

# varname,count,scope,agilo,agihi,fstatus,target
# XTOT,       0,    0,-9e99, 9e99,      0,  33e6
# e00300,     0,    1,-9e99, 9e99,      0,  20e9
# e00900,     0,    1,-9e99, 9e99,      0,  30e9
# e00200,     0,    1,-9e99, 9e99,      0,1000e9
# e02000,     0,    1,-9e99, 9e99,      0,  30e9
# e02400,     0,    1,-9e99, 9e99,      0,  60e9

```

## Get needed data

```{r}
#| label: get-soi-based-data

cdlong <- read_csv(fs::path(CDINTERMEDIATE, "cddata_long_clean.csv"))

```


## Create streamlined long CD-only file 

-   Drop non-CD records.
-   Drop variables we would never want.
-   Construct fstatus 0 records and records with counts by filing status.

### Drop non-CD records and variables we won't want

```{r}
#| label: drop-records-and-variables
#| output: false

cdlong1 <- cdlong |> 
  filter(rectype %in% c("cd", "cdstate", "DC"))

# quick data checks
cdlong1 |> filter(AGI_STUB==0) |> count(STATE)
cdlong1 |> filter(AGI_STUB==0) |> select(STATE, CONG_DISTRICT) |> distinct() |> nrow()
count(cdlong1, vtype)

# winnow data to variables we might consider targeting
# show candidates for dropping in a nice order
cdlong1 |> 
  filter(is.na(basevname)) |> 
  select(vname, description) |> 
  distinct() |> 
  arrange(vname)

dropvars <- c("CPREP", "DIR_DEP", "ELDERLY", "ELF", "PREP", "RAC", "TCE", 
              "TOTAL_VITA", "VITA", "VITA_EIC", "VRTCRIND")

cdlong2 <- cdlong1 |> 
  filter(!vname %in% dropvars)

rm(dropvars)

```


### Address agi issue

We have an issue with AGI: for Congressional Districts IRS does NOT report the number of returns with AGI. See the discussion on the introduction page.

We are going to create new rows for N00100 (Number of returns with AGI (estimated)), equal to N1, add it to the data.

```{r}
#| label: address-agi
#| output: false

nagi <- cdlong2 |> 
  filter(vname == "N1") |> 
  mutate(vname = "N00100",
         basevname = "v00100",
         description = "Number of returns with Adjusted Gross Income - AGI (estimated)")

cdlong3 <- bind_rows(cdlong2, nagi) |> 
  mutate(basevname = ifelse(vname == "A00100",
                            "v00100",
                            basevname))

# check
check <- cdlong3 |> 
  filter(basevname=="v00100", STATE=="WY") |> 
  arrange(STATE, CONG_DISTRICT, basevname, desc(vtype), AGI_STUB)

rm(nagi, check)

```


### Define statecd, fstatus, scope, count; put amounts in dollars; sort

```{r}
#| label: fstatus-misc
#| output: false

cdlong4 <- cdlong3 |> 
    mutate(
      fstatus = case_when(
        vname == "MARS1" ~ 1,
        vname == "MARS2" ~ 2,
        vname == "MARS4" ~ 4, # VERIFY WITH MARTIN
        .default = 0),
      
      count = case_when(
        vtype == "count" ~ 2, # correction - used to be 1
        vtype == "amount" ~ 0,
        .default = -99),
      
      scope = 1,
      value = ifelse(vtype == "amount",
                     value * 1000,
                     value)
      )
  
# summary(cdlong4)
# skim(cdlong4)
# count(cdlong4, fstatus)
# count(cdlong4, count, vtype)
# count(cdlong4, scope)

```

### Add Census population records to data

Prepare the Census population data.

```{r}
#| label: prepare-census-pop
#| output: false

# -   **varname**: XTOT
# -   **count**: 0
# -   **scope**: 0
# -   **agilo**: must be \< -8e99
# -   **agihi**: must be \> 8e99
# -   **fstatus**: 0
# -   **target**: area population

cdpop <- read_csv(fs::path(CDINTERMEDIATE, "cdpop1year.csv"))
glimpse(cdpop)

cdpop1 <- cdpop |> 
  select(STATEFIPS, STATE, CONG_DISTRICT, target=pop2021) |> 
  mutate(vname="XTOT",
         basevname="XTOT",
         description = "CD population in 2021 per Census ACS",
         AGI_STUB = 0,
         agirange = "Total",
         agilo = -9e99,
         agihi = 9e99,
         count = 0,
         scope = 0, 
         fstatus = 0)
  
# how well does cdpop merge against the soi data?
soistubs <- cdlong4 |> 
  select(STATEFIPS, STATE, CONG_DISTRICT, rectype, ndist) |> 
  distinct()

fmatch <- soistubs |> 
  left_join(cdpop1,
            by = join_by(STATEFIPS, STATE, CONG_DISTRICT))

fmatch |> filter(is.na(target)) # good, we matched on all CDs

# put rectype and ndist on the cdpop file
cdpop2 <- cdpop1 |> 
  left_join(fmatch |> 
              select(STATE, CONG_DISTRICT, rectype, ndist),
            by = join_by(STATE, CONG_DISTRICT))# statecd = paste0(STATE, CONG_DISTRICT),

rm(soistubs, fmatch)

```


```{r}
#| label: create-cdbasefile
#| output: false

cdbasefile <- bind_rows(cdlong4 |> 
                          rename(target=value) |> 
                          mutate(src="soi"),
                        cdpop2 |> mutate(src="census")) |> 
  mutate(statecd = paste0(STATE, CONG_DISTRICT),
         basevname=case_when(
           is.na(basevname) & vname=="A00101" ~ "v00101", # special handling
           is.na(basevname) ~ vname,
           .default = basevname)) |> # THINK ABOUT THIS CAREFULLY
  select(src, rectype, stabbr=STATE, cd=CONG_DISTRICT, statecd,
         agistub=AGI_STUB, agilo, agihi, basevname, scope, fstatus, count, target,
         vname, description, agirange) |> 
  arrange(statecd, src, scope, fstatus, basevname, count, agistub)
  
# glimpse(cdbasefile)
# summary(cdbasefile)
# skim(cdbasefile)

cdbasefile |> count(basevname)

cdbasefile |> filter(statecd=="WY00", agistub==0, basevname=="v00100")
cdbasefile |> filter(statecd=="WY00", agistub==0, basevname=="v00101")

write_csv(cdbasefile, fs::path(CDINTERMEDIATE, "cdbasefile_117.csv"))

```


